% Don't touch this %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage[left=1in,top=1in,right=1in,bottom=1in,headheight=3ex,headsep=3ex]{geometry}
\usepackage{graphicx}
\usepackage{float}
\usepackage{quoting}

\newcommand{\blankline}{\quad\pagebreak[2]}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Modify Course title, instructor name, semester here %%%%%%%%

\title{DATA 597: Automata}
\author{Calvin Deutschbein}
\date{Fall, 2025}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Don't touch this %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[sc]{mathpazo}
\linespread{1.05} % Palatino needs more leading (space between lines)
\usepackage[T1]{fontenc}
\usepackage[mmddyyyy]{datetime}% http://ctan.org/pkg/datetime
\usepackage{advdate}% http://ctan.org/pkg/advdate
\newdateformat{syldate}{\twodigit{\THEMONTH}/\twodigit{\THEDAY}}
\newsavebox{\MONDAY}\savebox{\MONDAY}{Mon}% Mon
\newcommand{\week}[1]{%
%  \cleardate{mydate}% Clear date
% \newdate{mydate}{\the\day}{\the\month}{\the\year}% Store date
  \paragraph*{\kern-2ex\quad #1, \syldate{\today} - \AdvanceDate[4]\syldate{\today}:}% Set heading  \quad #1
%  \setbox1=\hbox{\shortdayofweekname{\getdateday{mydate}}{\getdatemonth{mydate}}{\getdateyear{mydate}}}%
  \ifdim\wd1=\wd\MONDAY
    \AdvanceDate[7]
  \else
    \AdvanceDate[7]
  \fi%
}
\usepackage{setspace}
\usepackage{multicol}
%\usepackage{indentfirst}
\usepackage{fancyhdr,lastpage}
\usepackage{url}
\pagestyle{fancy}
\usepackage{hyperref}
\usepackage{lastpage}
\usepackage{amsmath}
\usepackage{layout}
\usepackage{csquotes}
\setlength\parindent{0pt}

\lhead{}
\chead{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Modify header here %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\rhead{\footnotesize Text in header}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Don't touch this %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\lfoot{}
\cfoot{\small \thepage/\pageref*{LastPage}}
\rfoot{}

\usepackage{array, xcolor}
\usepackage{color,hyperref}
\definecolor{clemsonorange}{HTML}{EA6A20}
\hypersetup{colorlinks,breaklinks,linkcolor=clemsonorange,urlcolor=clemsonorange,anchorcolor=clemsonorange,citecolor=black}

\begin{document}

\maketitle

\blankline

\begin{tabular*}{.93\textwidth}{@{\extracolsep{\fill}}lr}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Modify information %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
E-mail: \href{mailto:ckdeutschbein@willamette.edu}{\tt\bf ckdeutschbein@willamette.edu} & Web: \href{https://cd-public.github.io/compute/}{\tt\bf https://cd-public.github.io/compute/}  \\
\hline
\end{tabular*}

\vspace{5 mm}

% Zeroth Section %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Modality and Credit Hour Compliance}

\bigskip

\textbf{Student support} As a course intended to be completed with in person lecture with student questions, course TAs, and study groups, I will be available to meet weekly with the students, 1-on-1, to go over course material and homeworks.

\bigskip

\textbf{Asynchronous} I will upload a weekly lecture video to YouTube, with link posted to the course website on GitHub pages, every Monday from 8/25/25 to 11/17/25. Each lecture will cover the equivalent 4 credit hours of content, and is intended to be paused or replayed when students would typically ask questions in class. Together, this is 56 contact hours.

\bigskip

\textbf{Deliverables} Students will be responsible for developing a Quarto Book hosted on GitHub pages including results in LaTeX, Python and Graphviz for the theory of computation and the theory of complexity. Updates will be due every Monday at 12 midnight AOE. I will target 6-12 hours of effort each across weekly problem sets resulting in accordance with my understanding of credit hour policy.


% First Section %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Course Description}

Study of abstract models of computation, unsolvability, complexity theory, formal grammars and parsing, and other advanced topics in theoretical computer science.

% Second Section %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Course Materials}

\begin{itemize}
\item Course materials at \href{https://cd-public.github.io/compute/}{\tt\bf https://cd-public.github.io/compute/}
\item Optional Textbook: Sipser, Michael. Introduction to the Theory of Computation. 3rd ed. Cengage Learning, 2012. ISBN: 9781133187790
\item Supplemental Material: \href{https://ocw.mit.edu/courses/18-404j-theory-of-computation-fall-2020/pages/lecture-notes/}{\tt\bf Prof. Sipser's Lecture Notes}
\end{itemize}

% Third Section %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Prerequisite}
B.S. Computer Science or equivalent.

% Fourth Section %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Course Objectives}

\begin{enumerate}
    \item \textbf{Automata and Language Theory (18 contact hours)}
        \begin{itemize}
            \item \textbf{SLO 1:} Define and differentiate between finite automata (DFA, NFA, GNFA) and regular expressions, and demonstrate the equivalence between these models.
            \item \textbf{SLO 2:} Describe the capabilities and limitations of push-down automata (PDA) and context-free grammars (CFG), and apply the pumping lemma to prove that certain languages are not context-free.
        \end{itemize}
    \item \textbf{Computability Theory (18 contact hours)}
        \begin{itemize}
            \item \textbf{SLO 3:} Explain the concept of a Turing machine and its significance in computability theory.
            \item \textbf{SLO 4:} Discuss the Church-Turing thesis and its implications for the limits of computation.
            \item \textbf{SLO 5:} Define and distinguish between decidable and undecidable problems, and provide examples of each.
            \item \textbf{SLO 6:} Understand and apply concepts like the halting problem, reducibility, and the recursion theorem to analyze the computability of problems.
        \end{itemize}
    \item \textbf{Complexity Theory (18 contact hours)}
        \begin{itemize}
            \item \textbf{SLO 7:} Define and analyze the time and space complexity of algorithms using appropriate measures.
            \item \textbf{SLO 8:} Understand the significance of complexity classes P, NP, PSPACE, and NP-completeness, and discuss the implications of the P versus NP conjecture.
        \end{itemize}
\end{enumerate}

% 4.5th Section %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Assignments and Assessment}

\begin{itemize}
    \item Assignments in this course will have the sole purpose of student learning. Students will:
        \begin{itemize}
            \item  Begin the course with a grade of an "A".
            \begin{itemize}
                \item  Be expected to attend class.
                \item  Be expected to participate in class.
                \item  Be expected to complete assignments.
            \end{itemize}
            \item  Be contacted privately by the course instructor in the unusual event they are not meeting expectations.
            \begin{itemize}
                \item  Not have an expectation of perfection.
                \item  Not lose points or a grade without discussion.
                \item  Have a chance to explain their engagement with the course.
            \end{itemize}
            \item  Receive feedback collectively from the instructor and individually from peers.
            \begin{itemize}
                \item  Be entitled to individual feedback from the instructor at any time.
                \item  Receive narrative rather than quantitative feedback.
                \item  Receive positive and constructive feedback only.
            \end{itemize}
        \end{itemize}
\end{itemize}

\input{college}

\end{document}
