---
title: "Context Free Pumping"
author: Prof. Calvin
subtitle: "25 Jan 11:30 AM"
institute: CSCI 5100
format: 
    revealjs: 
        incremental: true
        theme: dark
        mainfont: monospace
        slide-number: true
        show-slide-number: all
        width: 1050
        height: 700
        footer: "[Home](../index.html)"
  
execute:
    echo: true
    cache: true
    freeze: true  # never re-render during project render

---

## Sketch

::: {.nonincremental}
- CF Pumping Lemma
  - Statement
  - Examples
:::

# Preface

## Avoid this trap

- Suppose we wish to prove a language is not a context free language
- We must prove there is no CFG/PDA that recognizes the language.
- It may be tempting to conclude:
  - I thought about it really hard.
  - I could find no PDA/CFG
  - Therefore the language is not a context free language.

## Example

- Take $\Sigma = \{0,1,2\}$
- Take $B = \{0^k1^k2^k| k \in \mathbb{N}\}$
- This language is not a context free language.
  - If you had a stack, match `0`'s with `1`'
  - How to deal with `2`'s?
  - Not a proof, but an intuition.

# Pumping Lemma

## Regular Pumping Lemma

$$
\begin{aligned}
&\forall A:\exists p \in \mathbb{N} : \\
&\exists xyz \in A : |xyz| \geq p \implies
\\
&\forall i \in \mathbb{N} : xy^iz \in A \land \\
&|y| > 0 \land \\
&|xy| \leq p
\end{aligned}
$$

- That is, $\{xz, xyz, xyyz\} \in A$
- We "pump up" the number of occurances of `y`

## Context Free Pumping Lemma

$$
\begin{aligned}
&\forall \text{ CFL } A :\exists p \in \mathbb{N} : \\
&\exists s = uvxyz \in A : |uvxyz| \geq p \implies
\\
&\forall i \in \mathbb{N} : uv^ixy^iz \in A \land \\
&|vy| > 0 \land \\
&|vxy| \leq p
\end{aligned}
$$

- That is, $\{uxz, uvxyz, uvvxyyz\} \in A$
- We "pump up" occurances of `v` and `y`

## Sketch of Proof

- We imagine an arbitrarily long string.
  - We'll be precise latter.
- The parse tree of a long string is very high.
  - Recall parse trees:

::: {.fragment}

```{dot}
// | echo: false
// | fig-height: 200px
digraph finite_automata {
  rankdir=TB; bgcolor="#191919";  
  node [fontcolor = "#ffffff", color = "#ffffff", fontsize="30"]
  edge [color = "#ffffff",fontcolor = "#ffffff"]

  node [shape=plaintext];
  1 [label="E"]; 

  2 [label="E"];
  3 [label="+"];
  4 [label="E"];

  5 [label="a"];
  6 [label="E"];
  7 [label="×"];
  8 [label="E"];


  1 -> 2
  1 -> 3
  1 -> 4

  2 -> 5
  4 -> 6
  4 -> 7
  4 -> 8

}
```

:::

## Checkin

- Can a string of length 1 million be made my a parse tree of height 1?
  - How about all multiple of 1 million?
- We'll quantify this shortly.

:::{.fragment}

$$
\begin{aligned}
    &S \rightarrow 0^{10^6} \\
\end{aligned}
$$

::: 

## Given a tall tree

- Let's assume the tall tree
- Each step in the tree involves variable substitution
- So as soon as height is greater than the size of set of variables, we necessarily repeat some variable.

## Tikz

- I'm going to use a new tool to show the next bit: Tikz!
  - Part of LaTeX
  - Seems to run well inside of [R](https://cran.r-project.org/web/packages/tikzDevice/vignettes/tikzDevice.pdf)
- Credit [LaTeX Graphics using TikZ](https://www.overleaf.com/learn/latex/LaTeX_Graphics_using_TikZ%3A_A_Tutorial_for_Beginners_(Part_1)%E2%80%94Basic_Drawing)
- Credit [JBGruber](https://stackoverflow.com/a/71856388)

## Example

```{r, engine = 'tikz'}
\begin{tikzpicture}
 \draw (0,0) circle (1cm);
\end{tikzpicture}
```

## Proof by Picture

:::{.r-stack}
```{r, engine = 'tikz'}
#| echo: false
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,6);
\draw[white] (1,2) -- node[below,cyan] {s} ++(8,0) ;
\end{tikzpicture}
```
:::


## $s$ is long

:::{.r-stack}
```{r, engine = 'tikz'}
#| echo: false
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,6);
\draw[white] (1,2) -- node[below,cyan] {s} ++(8,0) ;
\draw[orange, <->] (1,1) -- node[below,orange] {long} ++(8,0) ;

\end{tikzpicture}
```
:::

## Parse Tree

:::{.r-stack}
```{r, engine = 'tikz'}
#| echo: false
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,6);
\draw[white] (1,2) -- node[below,cyan] {s} ++(8,0) ;
\draw[orange, <->] (1,1) -- node[below,orange] {long} ++(8,0) ;
\draw[white] (1,2) -- (5,5) ;
\draw[white] (5,5) -- (9,2) ;

\end{tikzpicture}
```
:::

## Begin with $E$

:::{.r-stack}
```{r, engine = 'tikz'}
#| echo: false
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,6);
\draw[white] (1,2) -- node[below,cyan] {s} ++(8,0) ;
\draw[orange, <->] (1,1) -- node[below,orange] {long} ++(8,0) ;
\draw[white] (1,2) -- (5,5) ;
\draw[white] (5,5) -- (9,2) ;
\draw[white] (5,5.5) -- node[below,cyan] {E} ++(0,0)  ;

\end{tikzpicture}
```
:::


## Tree is tall

:::{.r-stack}
```{r, engine = 'tikz'}
#| echo: false
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,6);
\draw[white] (1,2) -- node[below,cyan] {s} ++(8,0) ;
\draw[orange, <->] (1,1) -- node[below,orange] {long} ++(8,0) ;
\draw[white] (1,2) -- (5,5) ;
\draw[white] (5,5) -- (9,2) ;
\draw[white] (5,5.5) -- node[below,cyan] {E} ++(0,0)  ;
\draw[orange, <->] (9,2) -- node[right,orange] {tall} ++(0,3) ;

\end{tikzpicture}
```
:::

## Take a Derivation

:::{.r-stack}
```{r, engine = 'tikz'}
#| echo: false
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,6);
\draw[white] (1,2) -- node[below,cyan] {s} ++(8,0) ;
\draw[orange, <->] (1,1) -- node[below,orange] {long} ++(8,0) ;
\draw[white] (1,2) -- (5,5) ;
\draw[white] (5,5) -- (9,2) ;
\draw[white] (5,5.5) -- node[below,cyan] {E} ++(0,0)  ;
\draw[orange, <->] (9,2) -- node[right,orange] {tall} ++(0,3) ;
\draw[white] (5,5) .. controls (4,4) and (6,3) .. (5,2);


\end{tikzpicture}
```
:::

## Necessary Repetition

:::{.r-stack}
```{r, engine = 'tikz'}
#| echo: false
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,6);
\draw[white] (1,2) -- node[below,cyan] {s} ++(8,0) ;
\draw[orange, <->] (1,1) -- node[below,orange] {long} ++(8,0) ;
\draw[white] (1,2) -- (5,5) ;
\draw[white] (5,5) -- (9,2) ;
\draw[white] (5,5.5) -- node[below,cyan] {E} ++(0,0)  ;
\draw[orange, <->] (9,2) -- node[right,orange] {tall} ++(0,3) ;
\draw[white] (5,5) .. controls (4,4) and (6,3) ..  (5,2);

\draw[white] (4.6,4.2) -- node[below,cyan] {R} ++(0,0) ;
\draw[white] (5.5,3.1) -- node[below,cyan] {R} ++(0,0) ;


\end{tikzpicture}
```
:::


## $R$ to $s$

:::{.r-stack}
```{r, engine = 'tikz'}
#| echo: false
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,6);
\draw[white] (1,2) -- node[below,cyan] {s} ++(8,0) ;
\draw[orange, <->] (1,1) -- node[below,orange] {long} ++(8,0) ;
\draw[white] (1,2) -- (5,5) ;
\draw[white] (5,5) -- (9,2) ;
\draw[white] (5,5.5) -- node[below,cyan] {E} ++(0,0)  ;
\draw[orange, <->] (9,2) -- node[right,orange] {tall} ++(0,3) ;
\draw[white] (5,5) .. controls (4,4) and (6,3) ..  (5,2);

\draw[white] (4.6,4.2) -- node[below,cyan] {R} ++(0,0) ;
\draw[white] (5.5,3.1) -- node[below,cyan] {R} ++(0,0) ;


\draw[white] (2.4,2) -- (4.85,3.8) ;
\draw[white] (4.85,3.8) -- (7,2) ;


\draw[white] (4.3,2) -- (5.25,2.8) ;
\draw[white] (5.25,2.8) -- (6.1,2) ;

\end{tikzpicture}
```
:::

## $s$ to $uvxyz$

:::{.r-stack}
```{r, engine = 'tikz'}
#| echo: false
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,6);
\draw[white] (1,2) -- node[below,cyan] {u} ++(1.6,0)  -- node[below,cyan] {v} ++(1.6,0) -- node[below,cyan] {x} ++(1.6,0) -- node[below,cyan] {y} ++(1.6,0) -- node[below,cyan] {z} ++(1.6,0);
\draw[orange, <->] (1,1) -- node[below,orange] {long} ++(8,0) ;
\draw[white] (1,2) -- (5,5) ;
\draw[white] (5,5) -- (9,2) ;
\draw[white] (5,5.5) -- node[below,cyan] {E} ++(0,0)  ;
\draw[orange, <->] (9,2) -- node[right,orange] {tall} ++(0,3) ;
\draw[white] (5,5) .. controls (4,4) and (6,3) ..  (5,2);

\draw[white] (4.6,4.2) -- node[below,cyan] {R} ++(0,0) ;
\draw[white] (5.5,3.1) -- node[below,cyan] {R} ++(0,0) ;


\draw[white] (2.4,2) -- (4.85,3.8) ;
\draw[white] (4.85,3.8) -- (7,2) ;


\draw[white] (4.3,2) -- (5.25,2.8) ;
\draw[white] (5.25,2.8) -- (6.1,2) ;

\end{tikzpicture}
```
:::

## Displace the lowest $R$...

:::{.r-stack}
```{r, engine = 'tikz'}
#| echo: false
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,6);
\draw[white] (1,2) -- node[below,cyan] {u} ++(1.6,0)  -- node[below,cyan] {v} ++(1.7,0) ;
\draw[white] (6.1,2)  -- node[below,cyan] {y} ++(1,0) -- node[below,cyan] {z} ++(1.9,0);
\draw[white] (1,2) -- (5,5) ;
\draw[white] (5,5) -- (9,2) ;
\draw[white] (5,5.5) -- node[below,cyan] {E} ++(0,0)  ;

\draw[white] (4.85,4.2) -- node[below,cyan] {R} ++(0,0) ;
\draw[white] (5.25,3.2) -- node[below,cyan] {R} ++(0,0) ;


\draw[white] (2.4,2) -- (4.85,3.8) ;
\draw[white] (4.85,3.8) -- (7,2) ;


\draw[white] (4.3,2) -- (5.25,2.8) ;
\draw[white] (5.25,2.8) -- (6.1,2) ;

\end{tikzpicture}
```
:::

## With the higher $R$...

:::{.r-stack}
```{r, engine = 'tikz'}
#| echo: false
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,6);
\draw[white] (1,2) -- node[below,cyan] {u} ++(1.6,0)  -- node[below,cyan] {v} ++(1.7,0) ;
\draw[white] (6.1,2)  -- node[below,cyan] {y} ++(1,0) -- node[below,cyan] {z} ++(1.9,0);
\draw[white] (1,2) -- (5,5) ;
\draw[white] (5,5) -- (9,2) ;
\draw[white] (5,5.5) -- node[below,cyan] {E} ++(0,0)  ;

\draw[white] (4.85,4.2) -- node[below,cyan] {R} ++(0,0) ;
\draw[white] (5.25,3.2) -- node[below,cyan] {R} ++(0,0) ;
\draw[white] (5.25,2.2) -- node[below,cyan] {R} ++(0,0) ;


\draw[white] (2.4,2) -- (4.85,3.8) ;
\draw[white] (4.85,3.8) -- (7,2) ;


\draw[white] (3.2,1) -- (5.25,2.8) ;
\draw[white] (5.25,2.8) -- (7.1,1) ;

\draw[white] (3.2,1) -- node[below,cyan] {v} ++(1.3,0) -- node[below,cyan] {x} ++(1.23,0) -- node[below,cyan] {y} ++(1.3,0) ;


\draw[white] (4.3,1) -- (5.25,1.8) ;
\draw[white] (5.25,1.8) -- (6.1,1) ;

\end{tikzpicture}
```
:::

## Side by side

::::{.columns}

:::{.column}

:::{.r-stack}
```{r, engine = 'tikz'}
#| echo: false
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,6);
\draw[white] (1,2) -- node[below,cyan] {u} ++(1.6,0)  -- node[below,cyan] {v} ++(1.6,0) -- node[below,cyan] {x} ++(1.6,0) -- node[below,cyan] {y} ++(1.6,0) -- node[below,cyan] {z} ++(1.6,0);
\draw[white] (1,2) -- (5,5) ;
\draw[white] (5,5) -- (9,2) ;
\draw[white] (5,5.5) -- node[below,cyan] {E} ++(0,0)  ;
\draw[white] (5,5) .. controls (4,4) and (6,3) ..  (5,2);

\draw[white] (4.6,4.2) -- node[below,cyan] {R} ++(0,0) ;
\draw[white] (5.5,3.1) -- node[below,cyan] {R} ++(0,0) ;


\draw[white] (2.4,2) -- (4.85,3.8) ;
\draw[white] (4.85,3.8) -- (7,2) ;


\draw[white] (4.3,2) -- (5.25,2.8) ;
\draw[white] (5.25,2.8) -- (6.1,2) ;

\end{tikzpicture}
```
:::

:::

:::{.column}

:::{.r-stack}
```{r, engine = 'tikz'}
#| echo: false
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,6);
\draw[white] (1,2) -- node[below,cyan] {u} ++(1.6,0)  -- node[below,cyan] {v} ++(1.7,0) ;
\draw[white] (6.1,2)  -- node[below,cyan] {y} ++(1,0) -- node[below,cyan] {z} ++(1.9,0);
\draw[white] (1,2) -- (5,5) ;
\draw[white] (5,5) -- (9,2) ;
\draw[white] (5,5.5) -- node[below,cyan] {E} ++(0,0)  ;

\draw[white] (4.85,4.2) -- node[below,cyan] {R} ++(0,0) ;
\draw[white] (5.25,3.2) -- node[below,cyan] {R} ++(0,0) ;
\draw[white] (5.25,2.2) -- node[below,cyan] {R} ++(0,0) ;


\draw[white] (2.4,2) -- (4.85,3.8) ;
\draw[white] (4.85,3.8) -- (7,2) ;


\draw[white] (3.2,1) -- (5.25,2.8) ;
\draw[white] (5.25,2.8) -- (7.1,1) ;

\draw[white] (3.2,1) -- node[below,cyan] {v} ++(1.3,0) -- node[below,cyan] {x} ++(1.23,0) -- node[below,cyan] {y} ++(1.3,0) ;


\draw[white] (4.3,1) -- (5.25,1.8) ;
\draw[white] (5.25,1.8) -- (6.1,1) ;

\end{tikzpicture}
```
:::

:::

::::

- On the left we have $s = uvxyz$
- On the right we have $s = uvvxyz = uv^2xy^2z$
- Process is repeatable.


## Deskew

::::{.columns}

:::{.column}

:::{.r-stack}
```{r, engine = 'tikz'}
#| echo: false
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,7);
\draw[white] (2,2) -- node[below,cyan] {u} ++(1,0)  -- node[below,cyan] {v} ++(1,0) ;
\draw[white] (4,2) -- node[below,cyan] {x} ++(2,0);
\draw[white] (6,2) -- node[below,cyan] {y} ++(1,0) -- node[below,cyan] {z} ++(1,0);

\draw[white] (5,5.5) -- node[below,cyan] {E} ++(0,0) ;

\draw[white] (5,4.5) -- node[below,cyan] {R} ++(0,0) ;
\draw[white] (5,3.5) -- node[below,cyan] {R} ++(0,0) ;

\draw[white] (2,2) -- (5,5) ;
\draw[white] (5,5) -- (8,2) ;

\draw[white] (3,2) -- (5,4) ;
\draw[white] (5,4) -- (7,2) ;

\draw[white] (4,2) -- (5,3) ;
\draw[white] (5,3) -- (6,2) ;

\end{tikzpicture}
```
:::

:::

:::{.column}

:::{.r-stack}
```{r, engine = 'tikz'}
#| echo: false
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,7);
\draw[white] (2,2) -- node[below,cyan] {u} ++(1,0)  -- node[below,cyan] {v} ++(1,0) ;
\draw[white] (6,2) -- node[below,cyan] {y} ++(1,0) -- node[below,cyan] {z} ++(1,0);

\draw[white] (5,5.5) -- node[below,cyan] {E} ++(0,0) ;

\draw[white] (5,4.5) -- node[below,cyan] {R} ++(0,0) ;
\draw[white] (5,3.5) -- node[below,cyan] {R} ++(0,0) ;
\draw[white] (5,2.5) -- node[below,cyan] {R} ++(0,0) ;

\draw[white] (2,2) -- (5,5) ;
\draw[white] (5,5) -- (8,2) ;

\draw[white] (3,2) -- (5,4) ;
\draw[white] (5,4) -- (7,2) ;

\draw[white] (3,1) -- (5,3) ;
\draw[white] (5,3) -- (7,1) ;

\draw[white] (4,1) -- (5,2) ;
\draw[white] (5,2) -- (6,1) ;

\draw[white] (3,1) -- node[below,cyan] {v} ++(1,0) -- node[below,cyan] {x} ++(2,0) -- node[below,cyan] {y} ++(1,0) ;

\end{tikzpicture}
```
:::

:::

::::

## Code Reveal

::::{.columns}

:::{.column}

```{.r}
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,7);
\draw[white] (2,2) -- node[below,cyan] {u} ++(1,0)  -- node[below,cyan] {v} ++(1,0) ;
\draw[white] (4,2) -- node[below,cyan] {x} ++(2,0);
\draw[white] (6,2) -- node[below,cyan] {y} ++(1,0) -- node[below,cyan] {z} ++(1,0);

\draw[white] (5,5.5) -- node[below,cyan] {E} ++(0,0) ;

\draw[white] (5,4.5) -- node[below,cyan] {R} ++(0,0) ;
\draw[white] (5,3.5) -- node[below,cyan] {R} ++(0,0) ;

\draw[white] (2,2) -- (5,5) ;
\draw[white] (5,5) -- (8,2) ;

\draw[white] (3,2) -- (5,4) ;
\draw[white] (5,4) -- (7,2) ;

\draw[white] (4,2) -- (5,3) ;
\draw[white] (5,3) -- (6,2) ;

\end{tikzpicture}
```

:::

:::{.column}

```{.r}
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,7);
\draw[white] (2,2) -- node[below,cyan] {u} ++(1,0)  -- node[below,cyan] {v} ++(1,0) ;
\draw[white] (6,2) -- node[below,cyan] {y} ++(1,0) -- node[below,cyan] {z} ++(1,0);

\draw[white] (5,5.5) -- node[below,cyan] {E} ++(0,0) ;

\draw[white] (5,4.5) -- node[below,cyan] {R} ++(0,0) ;
\draw[white] (5,3.5) -- node[below,cyan] {R} ++(0,0) ;
\draw[white] (5,2.5) -- node[below,cyan] {R} ++(0,0) ;

\draw[white] (2,2) -- (5,5) ;
\draw[white] (5,5) -- (8,2) ;

\draw[white] (3,2) -- (5,4) ;
\draw[white] (5,4) -- (7,2) ;

\draw[white] (3,1) -- (5,3) ;
\draw[white] (5,3) -- (7,1) ;

\draw[white] (4,1) -- (5,2) ;
\draw[white] (5,2) -- (6,1) ;

\draw[white] (3,1) -- node[below,cyan] {v} ++(1,0) -- node[below,cyan] {x} ++(2,0) -- node[below,cyan] {y} ++(1,0) ;

\end{tikzpicture}
```

:::

::::

## $uxz$

::::{.columns}

:::{.column}

:::{.r-stack}
```{r, engine = 'tikz'}
#| echo: false
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,7);
\draw[white] (2,2) -- node[below,cyan] {u} ++(1,0) ;
\draw[white] (4,3) -- node[below,cyan] {x} ++(2,0);
\draw[white] (7,2) --  node[below,cyan] {z} ++(1,0);

\draw[white] (5,5.5) -- node[below,cyan] {E} ++(0,0) ;

\draw[white] (5,4.5) -- node[below,cyan] {R} ++(0,0) ;

\draw[white] (2,2) -- (5,5) ;
\draw[white] (5,5) -- (8,2) ;

\draw[white] (3,2) -- (5,4) ;
\draw[white] (5,4) -- (7,2) ;


\end{tikzpicture}
```
:::

:::

:::{.column}

:::{.r-stack}
```{r, engine = 'tikz'}
#| echo: false
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,7);
\draw[white] (2,2) -- node[below,cyan] {u} ++(1,0)  -- node[below,cyan] {v} ++(1,0) ;
\draw[white] (6,2) -- node[below,cyan] {y} ++(1,0) -- node[below,cyan] {z} ++(1,0);

\draw[white] (5,5.5) -- node[below,cyan] {E} ++(0,0) ;

\draw[white] (5,4.5) -- node[below,cyan] {R} ++(0,0) ;
\draw[white] (5,3.5) -- node[below,cyan] {R} ++(0,0) ;
\draw[white] (5,2.5) -- node[below,cyan] {R} ++(0,0) ;

\draw[white] (2,2) -- (5,5) ;
\draw[white] (5,5) -- (8,2) ;

\draw[white] (3,2) -- (5,4) ;
\draw[white] (5,4) -- (7,2) ;

\draw[white] (3,1) -- (5,3) ;
\draw[white] (5,3) -- (7,1) ;

\draw[white] (4,1) -- (5,2) ;
\draw[white] (5,2) -- (6,1) ;

\draw[white] (3,1) -- node[below,cyan] {v} ++(1,0) -- node[below,cyan] {x} ++(2,0) -- node[below,cyan] {y} ++(1,0) ;

\end{tikzpicture}
```
:::

:::

::::

## Context Free Pumping Lemma

::::{.columns}

:::{.column}

$$
\begin{aligned}
&\forall \text{ CFL } A :\exists p \in \mathbb{N} : \\
&\exists s = uvxyz \in A : |uvxyz| \geq p \implies
\\
&\forall i \in \mathbb{N} : uv^ixy^iz \in A \land \\
&|vy| > 0 \land \\
&|vxy| \leq p
\end{aligned}
$$


:::

:::{.column}

*Proof*

:::{.r-stack}
```{r, engine = 'tikz'}
#| echo: false
\begin{tikzpicture}
\fill[darkgray!40!black] (0,0) rectangle (10,7);
\draw[white] (2,2) -- node[below,cyan] {u} ++(1,0)  -- node[below,cyan] {v} ++(1,0) ;
\draw[white] (6,2) -- node[below,cyan] {y} ++(1,0) -- node[below,cyan] {z} ++(1,0);

\draw[white] (5,5.5) -- node[below,cyan] {E} ++(0,0) ;

\draw[white] (5,4.5) -- node[below,cyan] {R} ++(0,0) ;
\draw[white] (5,3.5) -- node[below,cyan] {R} ++(0,0) ;
\draw[white] (5,2.5) -- node[below,cyan] {R} ++(0,0) ;

\draw[white] (2,2) -- (5,5) ;
\draw[white] (5,5) -- (8,2) ;

\draw[white] (3,2) -- (5,4) ;
\draw[white] (5,4) -- (7,2) ;

\draw[white] (3,1) -- (5,3) ;
\draw[white] (5,3) -- (7,1) ;

\draw[white] (4,1) -- (5,2) ;
\draw[white] (5,2) -- (6,1) ;

\draw[white] (3,1) -- node[below,cyan] {v} ++(1,0) -- node[below,cyan] {x} ++(2,0) -- node[below,cyan] {y} ++(1,0) ;

\end{tikzpicture}
```
:::

:::

::::

## Details

- Take $b$ maximum branch size
- Take $h$ height parse tree
- Need $|s| < b^h$
- Let $p = b^{|V|}$
  - $V$ is the set of variables.

## All conditions

1. $uv^ixy^iz \in A \forall i \in \mathbb{N}$
  -  Cut and paste
2. $vy \neq \varepsilon$
  - Take smallest possible tree.
3. $|vxy| \leq p$
  - Pick lowest $R$

## Example

- Take $B = \{0^k1^k2^k|k\in\mathbb{N}\}$
  - Need `0`, `1`, and `2` in $vxy$
    - $|vxy| \leq p$
  - Pumping anything leads to unequal.
  - Not in CFL by condition 3.

<!-- https://www.youtube.com/watch?v=IycOPFmEQk8&t=2311s -->

# Stretch Break

- [Home](https://cd-public.github.io/compute/)